---
title: "Worksheet#4C"
author: "Marivic Amuan"
date: "2023-11-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#1. Use the dataset mpg

#1A. Show your solutions on how to import a csv file into the environment.

library(readr)

mpg <- read_csv("mpg.csv")

head(mpg)

```

```{r}

#1B. Which variables from mpg dataset are categorical?

str(mpg)

#ANSWER: Manufacturer,model,year,cyl,trans,drv,fl, and class are the categorical variables.
```

```{r}
#1C. Which are continuous variables?
#The continious variablesa are the , manufacturer, model, disply, year , cyl, cty ,hwy, fl, trans, drv and class.

summary(mpg)

```

```{r}
#2. Which manufacturer has the most models in this data set? Which model has the most variations? Show your answer.

#ANSWER: THE MOST MODEL MANUFACTURER IS DODGE.THE MOST VARIATION IS CARAVAN 2WD.
#ANSWER: #The model with most variation is caravan 2wd.
library(magrittr)
library(dplyr)

model <- mpg %>%
 group_by(manufacturer) %>%
 summarise(count = n()) %>%
 arrange(desc(count))

print(model)

count <- mpg %>%
 group_by(model) %>%
 summarise(variation = n()) %>%
 arrange(desc(variation))

print(count)

```


```{r}
#2A Group the manufacturers and find the unique models. Show your codes and result.
library(dplyr)

manumodel <- mpg %>%
 group_by(manufacturer) %>%
 summarise(unique_models = n_distinct(model)) 

print(manumodel)
```


```{r}
#2B. Graph the result by using plot() and ggplot(). Write the codes and its result.

library(ggplot2)

plot(ggplot(manumodel, aes(x = reorder(manufacturer, -unique_models), y = unique_models, fill = unique_models)) +
 geom_bar(stat = "identity", width = 0.8) + 
 labs(title = "Number of Manufacturer's Unique Models", 
      x = "Manufacturer",
      y = "Number of Unique Models") +
   
 theme_minimal() +
 scale_fill_gradient(low = "black", high = "pink") +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))) 
```

```{r}
#2.2 A What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?

#This plot displays the relationship between car models on the x-axis and manufacturers (on the y-axis) using individual points.

 ggplot(mpg, aes(model, manufacturer)) + geom_point() 

```

```{r}
#2B.b. For you, is it useful? If not, how could you modify the data to make it more informative?

#To enhance its informativeness, I plan to introduce color coding to distinguish points according to various variables. Additionally, I'll include a legend to aid viewers in interpreting the data without confusion. To improve readability, I'll adjust the size and angle of variable names.

ggplot(mpg, aes(x = model, y = manufacturer, color = class)) + 
  geom_point() +
  labs(title = "Car Models and Manufacturers",
       cex = 3,
       x = "Car Model",
       y = "Manufacturer",
       color = "Vehicle Class") +
  theme_minimal() +
  theme(legend.position = "right", axis.text.x = element_text(angle = 90, hjust = 1),
  plot.title = element_text(hjust  = 0.6))
```

```{r}
#3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.

library(ggplot2)
library(dplyr)

data(mpg)


mean_displ_df <- mpg %>%
 group_by(year, model) %>%
 summarise(mean_displ = mean(displ)) %>%
 arrange(desc(mean_displ)) %>%
 filter(row_number() < 20) 

plot <- ggplot(mean_displ_df, aes(x = year, y = mean_displ, fill = model)) +
 geom_bar(stat = "identity", position = "dodge") +
 theme_minimal() +
 guides(fill = guide_legend(ncol = 2)) 
 labs(title = "Average Engine Displacement over the years for the top 20 models",
       x = "Year",
       y = "Engine Displacement",
       fill = "Model")

print(plot)


```

```{r}
#4.  Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result


library(dplyr)
data(mpg)


countmodel <- mpg %>%
  group_by(model) %>%
  summarise(num_cars = n())


print(countmodel)

```

```{r}

#4A. Plot using geom_bar() using the top 20 observations only. The graphs should have a title, labels and colors. Show code and results.


library(ggplot2)
library(dplyr)

data(mpg)

summary_data <- mpg %>%
  count(model) %>%
  arrange(desc(n)) %>%
  slice(1:20)

top_models <- summary_data$model
palette <- scales::hue_pal()(length(top_models))

summary_data <- summary_data %>%
  mutate(color = palette[match(model, top_models)])

ggplot(summary_data, aes(x = reorder(model, n), y = n, fill = model)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Top 20 Car Models by Count",
    x = "Car Models",
    y = "Total Cars"
  ) +
  scale_fill_manual(values = palette, name = "Car Models", breaks = summary_data$model) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.key.size = unit(0.1, "cm"),
    plot.title = element_text(hjust = 0.5)  
  )


```

```{r}
#4B. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.


library(ggplot2)
library(dplyr)

data(mpg)

summary_data <- mpg %>%
  count(model) %>%
  arrange(desc(n)) %>%
  slice(1:20)


top_models <- summary_data$model
palette <- scales::hue_pal()(length(top_models))

summary_data <- summary_data %>%
  mutate(color = palette[match(model, top_models)])

ggplot(summary_data, aes(x = reorder(model, n), y = n, fill = model)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Top 20 Car Models by Count",
    y = "Car Models",  
    x = "Total Cars"
  ) +
  scale_fill_manual(values = palette, name = "Car Models", breaks = summary_data$model) +
  coord_flip() + 
  theme_minimal() +
  theme(
    legend.key.size = unit(0.1, "cm"),
    plot.title = element_text(hjust = 0.5)  
  )

```

```{r}
#5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using#geom_point with aesthetic color = engine displacement. Title should be “Relationship between

#No. of Cylinders and Engine Displacement”.

library(ggplot2)
library(dplyr)

data(mpg)

ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +  
  labs(
    title = "Relationship between No. of Cylinders and Engine Displacement",
    x = "Number of Cylinders",
    y = "Engine Displacement")

```

```{r}
#5A. How would you describe its relationship? Show the codes and its result.


#Utilizing a regression line, showcase the correlation between the number of cylinders and engine displacement. 
#As the cylinder count rises, there's a tendency for the engine size to also increase.

library(ggplot2)
library(dplyr)

data(mpg)

ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +  
  labs(
    title = "Relationship between No. of Cylinders and Engine Displacement",
    x = "Number of Cylinders",
    y = "Engine Displacement"
  )

```

```{r}
#6. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
#it with a continuous variable you have identified in #1-c. What is its result? Why it produced such
#output? 

library(ggplot2)
library(dplyr)

data(mpg)

ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point() +
  labs(
    title = "Relationship between Engine Displacement and Highway MPG",
    x = "Engine Displacement",
    y = "Highway MPG"
  ) +
  scale_color_gradient(low = "black", high = "pink")  


```
The scatter plot illustrates the relationship between engine displacement (displ) and highway miles per gallon (hwy). It employs a color gradient derived from city miles per gallon (cty) to represent a continuous variable across the plotted points. The color gradient, reflective of city miles per gallon (cty), doesn't directly correlate with engine displacement or highway miles per gallon (displ and hwy). Rather, it showcases the diversity in city MPG across the entire scatter plot.

```{r}
#6. Import the traffic.csv onto your R environment.
#A. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
#your answer.

traffic <- read_csv("traffic.csv")
head(traffic)


observations <- nrow(traffic)
variables <- names(traffic)

cat("Number of observations:", observations, "\n")
cat("The variables are:", variables, "\n")


```

```{r}
#6B. subset the traffic dataset into junctions. What is the R codes and its output?

junctions1 <- subset(traffic, Junction == 1)
junctions2 <- subset(traffic, Junction == 2)
junctions3 <- subset(traffic, Junction == 3)
junctions4 <- subset(traffic, Junction == 4) 

#The output are:
junctions1  
junctions2
junctions3
junctions4

```


```{r}
#6C. Plot each junction in a using geom_line(). Show your solution and output.

#Junction 1

ggplot(junctions1, aes(x = DateTime, y = Vehicles, color = "Junction 1")) +
  geom_line() +
  labs(
    title = "Traffic Volume at Junction 1",
    x = "Date",
    y = "Volume"
  ) +
  scale_color_discrete(name = "Junction") +
 theme(plot.title = element_text(hjust = 0.5))

#Junction 2

ggplot(junctions2, aes(x = DateTime, y = Vehicles, color = "Junction 2")) +
  geom_line() +
  labs(
    title = "Traffic Volume at Junction 2",
    x = "Date",
    y = "Volume"
  ) +
  scale_color_discrete(name = "Junction") +
  theme(plot.title = element_text(hjust = 0.5))

#Junction 3

ggplot(junctions3, aes(x = DateTime, y = Vehicles, color = "Junction 3")) +
  geom_line() +
  labs(
    title = "Traffic Volume at Junction 3",
    x = "Date",
    y = "Volume"
  ) +
  scale_color_discrete(name = "Junction") +
 theme(plot.title = element_text(hjust = 0.5))

#Junction 4

ggplot(junctions4, aes(x = DateTime, y = Vehicles, color = "Junction 4")) +
  geom_line() +
  labs(
    title = "Traffic Volume at Junction 4",
    x = "Date",
    y = "Volume"
  ) +
  scale_color_discrete(name = "Junction") +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}
#7. From alexa_file.xlsx, import it to your environment

library(readxl)
alexa_file <- read_excel("alexa_file.xlsx")
head(alexa_file)

#7A. How many observations does alexa_file has? What about the number of columns? Show your solution and answer. 

observations <- nrow(alexa_file)
columns <- ncol(alexa_file)

cat("Number of observations:", observations, "\n")
cat("Number of columns:", columns, "\n")

#The number of observations is 3,150 and The number of columns is 5. 
```

```{r}
#7B. group the variations and get the total of each variations. Use dplyr package. Show solution and answer.

library(dplyr)

result <- alexa_file %>%
  group_by(variation) %>%
  summarise(total_variations = n()) 

print(result)

```

```{r}
#7C. Plot the variations using the ggplot() function. What did you observe? Complete the details of the graph. Show solution and answer. 


library(ggplot2)

var <- ggplot(result, aes(x = variation, y = total_variations, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Variations Count",
       x = "Variation",
       y = "Total Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_discrete(name = "Variation") +
  guides(fill = guide_legend(ncol = 2)) +
  theme(plot.title = element_text(hjust = 0.5)) 

print(var)

```
This visualization represents the different variations of the Alexa File, utilizing distinct colors to label each variation, aiding viewers in analysis. Additionally, it includes the total count for each variation. The variation labeled "Black Dot" notably appears more frequently than the others. The legend, conveniently organized into two columns, provides a clear reference to associate each color with its respective type of variation.



```{r}


#7D. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the graphs. Show your answer and solution.


library(dplyr)
library(ggplot2)

alexa_file$date <- as.Date(alexa_file$date)
alexa_file$month <- format(alexa_file$date, "%m")

monthcount <- alexa_file %>%
   count(month)

 p <- ggplot(monthcount, aes(x = as.integer(month), y = n, color = "Reviews")) +
  geom_line(size = 1) +
  labs(title = "Number of Verified Reviews Over Time",
       x = "Month",
       y = "Number of Verified Reviews",
       color = "Legend Title") +  # Change legend title
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  scale_color_manual(values = c("blue"), labels = c("Reviews")) +  
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

print(p) 

```

```{r}

#7E. Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a #relationship. Show your solution and answer.

library(dplyr)
library(ggplot2)

variation_ratings <- alexa_file %>%
  group_by(variation) %>%
  summarize(avg_rating = mean(rating))
print(variation_ratings)

highest <- variation_ratings %>% 
  filter(avg_rating == max(avg_rating))
print(highest) 

variation_names <- variation_ratings$variation
average_ratings <- variation_ratings$avg_rating

barplot(average_ratings, names.arg = variation_names, col = "green",
        main = "Average Rating by Variation",
        xlab = "Variation", ylab = "Average Rating",
        cex.axis = 0.8, cex.names = 0.5, las = 2)  

top_variation <- variation_names[which.max(average_ratings)]
top_rating <- max(average_ratings)

cat("The variation with the highest average rating is:", top_variation, "with an average rating of", top_rating)

```